Should I Airbnb My Property?

### 2017-03-08 abb_Writeup.Rmd
find and replace terms

  source(paste0(code.path, "dataviztools/ggplottools.R"))
  source(paste0(code.path, "datamgmttools/dataMungetools.R"))
  source(paste0(code.path, "research/AirBNBMelbourne/abb_Functions.R"))

@@

&nbsp;
&nbsp;

@@

We begin by filtering observations by time. Our hypothetical leasing strategy decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long-term rental versus an Airbnb-type, short-term rental approach to generating income. To determine the likely LTR rate for any given property, we will use the observed rental transactions from September 1, 2014 to August 31, 2015. Taking the Airbnb or STR approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors). In this simulated example we will use AirBnB data from September 1, 2015 to August 31, 2016 to represent the actual AirBNB market conditions over the time period in question.

Next, we filter the data based on property types. AirBNB units are classified into 19 different property types. The most common are Apartments and Houses, the least common are Igloo, Tent, Treehouse and Yurt.^[Property types are owner/lister defined, therfore it is likely that some designation, such as Igloo, at the very least, are not completely true representations of the property type] For the sake of this analysis, we will collapse these twenty types into three types:

@@

Within the long-term rental data, property types fall into eight categories.  Like the AirBNB data we group these into three categories matching as best as possible to the AirBNB categories:

@@

&nbsp;

@@

As the distribution of occupancy rates is not uniform the initial plot may be misleading as to the frequency of short-term rentals being more profitable than long term. In the right hand panel of \autoref{occpref} (right panel) we transform the x-axis to represent the quantile of the full distribution of occupancy rates. By doing so, we see that, in general, properties need to be in the upper 25th percentile of occupancy rates across the entire market in order to have a equal likelihood of being more profitable as a short-term rental than as a more traditional long term one.

&nbsp;

@@

  qtlocc <- ggplot(qtlocc$data,
                   aes(x=x.var, y=Var))+
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate (Quantile)') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25th', '50th', '75th', '100th')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='none') + 
    ggtitle('Quantile Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

@@

```{r, fig.width=9, fig.height=4, message=FALSE, warning=FALSE, echo=FALSE, fig.cap='\\label{smocc}Analysis of Quantile Rates: Metro Level'}

@@

  smq.occ <- ggplot(smqocc$data,
                    aes(x=x.var, y=Var, group=ID, color=ID)) +  
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate (Quantile)') +
    scale_color_manual(values=sm.col,
                       name='') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25th', '50th', '75th', '100th')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='bottom') + 
    ggtitle('Quantile Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

@@

&nbsp;

```{r, fig.width=8, fig.height=4, echo=FALSE, fig.cap='Analysis of Quantile Rates: Metro Level'}

@@

 ## Combine into full table of coef and diags
  #full.table[,2] <- gsub(" ", "&nbsp;", full.table[,2], fixed=TRUE)

@@

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.cap='\\label{pothm}Heatmaps of Potential Short Term Preference'}

@@

```{r build_blockrate, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE}

  block.rate.period <- ggplot(str.data, aes(x=block.rate, y=nbr.block*extr.par)) + 
  geom_point(fill=str.col, color=str.col, size=.2) +
  ylab('Nbr of Blocked Periods\n') + 
  xlab('\n% of Time Property is Blocked') +
  scale_x_continuous(breaks=seq(0, 1, .2),
                     labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +
    theme(plot.title = element_text(hjust = 0.5))

@@

```{r plot_hostblock, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3.5, fig.caption='\\label{hostblock}Host Blocking Tendencies', fig.lp='hostblock'}

@@

```{r plot_hosttypes, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3.5, fig.caption='\\label{hostcat}Host Categories', fig.lp='hosttypes'}

@@

```{r plot_hostbar, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3, fig.width=4, fig.caption='\\label{htc}Host Type Counts'}

host.type.bar <- ggplot(str.data, aes(host.type, fill=host.type)) +
  geom_bar() +
  scale_fill_manual(values=c(abb.col[2], abb.col[4], abb.col[7], abb.col[9]))+
  theme(legend.position='none') +
  ylab('Frequency\n') + 
  xlab('\nHost Type') +
  #ggtitle('Short Term Host Types') + 
  theme(plot.title = element_text(hjust = 0.5))

>>>>

  source(paste0(code.path, "dataviztools/ggplottools.R"))
  source(paste0(code.path, "datamgmttools/dataMungetools.R"))
  source(paste0(code.path, "research/AirbnbMelbourne/abb_Functions.R"))

@@

@@

We begin by filtering observations by time. Our hypothetical leasing strategy decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long-term rental versus an Airbnb-type, short-term rental approach to generating income. To determine the likely LTR rate for any given property, we will use the observed rental transactions from September 1, 2014 to August 31, 2015. Taking the Airbnb or STR approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors). In this simulated example we will use Airbnb data from September 1, 2015 to August 31, 2016 to represent the actual Airbnb market conditions over the time period in question.

Next, we filter the data based on property types. Airbnb units are classified into 19 different property types. The most common are Apartments and Houses, the least common are Igloo, Tent, Treehouse and Yurt.^[Property types are owner/lister defined, therfore it is likely that some designation, such as Igloo, at the very least, are not completely true representations of the property type] For the sake of this analysis, we will collapse these twenty types into three types:

@@

Within the long-term rental data, property types fall into eight categories.  Like the Airbnb data we group these into three categories matching as best as possible to the Airbnb categories:

@@

@@

As the distribution of occupancy rates is not uniform the initial plot may be misleading as to the frequency of short-term rentals being more profitable than long term. In the right hand panel of \autoref{occpref} (right panel) we transform the x-axis to represent the percentile of the full distribution of occupancy rates. By doing so, we see that, in general, properties need to be in the upper 25th percentile of occupancy rates across the entire market in order to have a equal likelihood of being more profitable as a short-term rental than as a more traditional long term one.

@@

  qtlocc <- ggplot(qtlocc$data,
                   aes(x=x.var, y=Var))+
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate (Pcntl)') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25th', '50th', '75th', '100th')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='none') + 
    ggtitle('Percentile Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

@@

```{r, fig.width=9, fig.height=4, message=FALSE, warning=FALSE, echo=FALSE, fig.cap='\\label{smocc}Analysis of Percentile Rates: Metro Level'}

@@

  smq.occ <- ggplot(smqocc$data,
                    aes(x=x.var, y=Var, group=ID, color=ID)) +  
    geom_point(alpha=.5, size=.4) +
    stat_smooth(se=FALSE, size=2, span=.5) +
    xlab('\nOccupancy Rate (Pcntl)') +
    scale_color_manual(values=sm.col,
                       name='') +
    scale_x_continuous(breaks=seq(0, 100, by=25),
                       labels=c('0%', '25th', '50th', '75th', '100th')) +
    ylab('\nSTR Preferability %') +
    scale_y_continuous(breaks=seq(0, 1, by=.25),
                       labels=c('0%', '25%', '50%', '75%', '100%')) +
    theme(legend.position='bottom') + 
    ggtitle('Percentile Rates') + 
    theme(plot.title = element_text(hjust = 0.5))

@@

```{r, fig.width=8, fig.height=4, echo=FALSE, fig.cap='Analysis of Percentile Rates: Metro Level'}

@@

 ## Combine into full table of coef and diags
  #full.table[,2] <- gsub(" ", "", full.table[,2], fixed=TRUE)

@@

```{r, results='asis', echo=FALSE, message=FALSE, comment=FALSE, fig.height=3.5, fig.cap='\\label{pothm}Heatmaps of Potential Short Term Preference'}

@@

```{r build_blockrate, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE}

  block.rate.period <- ggplot(str.data, aes(x=block.rate, y=nbr.block*extr.par)) + 
  geom_point(fill=str.col, color=str.col, size=.2) +
  ylab('Nbr of Blocked Periods\n') + 
  xlab('\n% of Time Property is Blocked') +
  scale_x_continuous(breaks=seq(0, 1, .2),
                     labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    #ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +
    theme(plot.title = element_text(hjust = 0.5))

@@

```{r plot_hostblock, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.height=3.5, fig.cap='\\label{hostblock}Host Blocking Tendencies'}

@@

```{r plot_hosttypes, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, fig.height=3.5, fig.cap='\\label{hostcat}Host Categories', fig.lp='hosttypes'}

@@

```{r plot_hostbar, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.height=3, fig.width=4, fig.cap='\\label{htc}Host Type Counts'}

host.type.bar <- ggplot(str.data, aes(host.type, fill=host.type)) +
  geom_bar() +
  scale_fill_manual(values=c(abb.col[2], abb.col[4], abb.col[7], abb.col[9]))+
  theme(legend.position='none') +
  ylab('Frequency\n') + 
  xlab('\nHost Type') +
  scale_x_discrete(labels=c('Profit','Sharer','Multi','Unk'))+
  #ggtitle('Short Term Host Types') + 
  theme(plot.title = element_text(hjust = 0.5))
